{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-08-14 thesis/landslide.tex
test case misbehave (is dumb)

Many of the supplied P2 test cases invoke the {\tt misbehave()} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff (\sect{\ref{sec:410-secrecy}}).
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

To illustrate, consider the following program.

% TODO
% TODO - make a progr that data races immediately after thr create, no mutex es or whatever.
% TODO - check if it rly happens under pure happens before? or if this is only a concern under limited HB.

>>>>

Many of the supplied P2 test cases invoke the {\tt misbehave} system call
with a mysterious argument (usually {\tt BGND\_BRWN >> FGND\_CYAN})
before the creation of any child threads.
The use of terminal color code constants is of course a red herring of obfuscation,
as the true nature of the Pathos reference kernel's {\tt misbehave} modes
is a closely-guarded secret among 15-410 course staff (\sect{\ref{sec:410-secrecy}}).
The mode in question causes the reference kernel to prioritize scheduling the child thread over the parent
whenever {\tt thread\_fork} is called,
and the target thread over the invoking thread
whenever {\tt make\_runnable} is called,
which are necessary to allow Landslide to recognize a {\tt yield()} preemption point
and be able to run the newly-runnable thread as soon as possible.

To illustrate, consider the following program in Figure~\ref{fig:misbehave-example},
and suppose Landslide is configured to preempt only on mutex API calls
(such as in the first step of Iterative Deepening (\sect{\ref{sec:quicksand-id}})).
Because Landslide ignores all kernel-level synchronization short of context switches when testing user-level code,
if the kernel created the child thread and returned from {\tt thread\_fork}
(the system call underlying {\tt thr\_create()})
without yielding first,
the next preemption point will not occur until {\tt thr\_join()} waits for the child to exit.
Hence, DPOR will erroneously think everything before that {\tt thr\_join()}
happens-before (\sect{\ref{sec:landslide-dpor-hb}}) anything the child does,
and will fail to identify the racing accesses on {\tt x}.

\begin{figure}[h]
	\begin{center}
	\begin{tabular}{l}
		\texttt{\ctype{void} \call{child}(\ctype{int} *xp) \{} \\
		\texttt{~~~~*xp++;} \\
		\texttt{\}} \\
		\\
		\texttt{\ctype{void} \call{parent}() \{} \\
		\texttt{~~~~\ctype{int} x = \const{0};} \\
		\texttt{~~~~\ctype{int} tid = \call{thr\_create}(child, \&x);} \\
		\texttt{~~~~x++;} \\
		\texttt{~~~~\call{thr\_join}(tid, \const{NULL});} \\
		\texttt{~~~~\flow{assert}{}(x == \const{2});} \\
		\texttt{\}} \\
	\end{tabular}
	\end{center}
	\caption{Example demonstrating the need for {\tt misbehave} to force the kernel
	to yield during {\tt thread\_fork}.}
	\label{fig:misbehave-example}
\end{figure}

Though Iterative Deepening's soundness (\sect{\ref{sec:quicksand-soundness}})
guarantees all data races will eventually be detected starting from just synchronization preemption points,
it assumes threads becoming runnable counts among those.
In that sense, {\tt misbehave} serves to restore the last synchronization preemptions where they belong.
If at this point the reader wonders why Landslide doesn't just identify
the {\tt thread\_create} and {\tt make\_runnable} system calls in the arbiter itself (\sect{\ref{sec:landslide-arbiter}})
and skip this mysterious user-visible complexity,
they would be right to ask:
I have left it this way for no better reason than to maintain consistency
with the upcoming chapters' experimental environments,
and intend on fixing it in a future update.
